% Suggested Viewer : pympress
\documentclass[xcolor=dvipsnames]{beamer}

\usepackage[utf8]{inputenc} 
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
\usepackage{multimedia}

\usetheme{Warsaw}

% Enable the use of notes on second screen
\usepackage{pgfpages}
% \setbeameroption{show notes on second screen}
%\graphicspath{...}

\input{style.tex}
\input{headpage.tex}


\title{Roadmap}

\subtitle{Current status and upcoming changes}

\author[L. Hofer \& M. Paetzel]{Ludovic Hofer\\Maike Paetzel}

\newcommand{\presentationdate}{
  June 26th 2020
}

\institute{Institute}

\begin{document}

%TODO add link on first page?
\begin{frame}[plain]
  \maketitle
\end{frame}
\note{TODO}

% \begin{frame}[plain]
%   \tableofcontents
% \end{frame}

\section{Introduction}

\begin{frame}{Motivations for the update}
  \begin{itemize}
  \item Make link between scientific challenges and roadmap more explicit
  \item Ensure that rules evolve along with robot capabilities
  \item Allow teams to plan their development 
  \item Create a maintainable document
  \end{itemize}
\end{frame}

\begin{frame}{Structure of the new roadmap}
  \begin{block}{Scientific challenges}
    \begin{itemize}
    \item Scientific context of the league
    \item Research contributions from the league
    \item Require a major update every 4 years
    \end{itemize}
  \end{block}
  \begin{block}{Short-Term: Upcoming changes}
    \begin{itemize}
    \item Changes planned for the next 5 years
    \end{itemize}
  \end{block}
  \begin{block}{Long-term: event-triggered roadmap}
    \begin{itemize}
    \item What are the remaining steps to win against humans?
    \item Metrics: how are improvements and performance measured
    \item Criteria required to take significant steps
    \end{itemize}
  \end{block}
\end{frame}

\section{Short term: upcoming changes}

\begin{frame}{Implemented for 2020}
  \begin{itemize}
  \item Structural changes to the league
    \begin{itemize}
    \item TeenSize now plays in Kid and Adult
    \item Division A/B
    \item Combined teams
    \end{itemize}
  \item Autonomous placement for all robots
  \item Humanoid Research Demonstration
  \item Technical Challenges
    \begin{itemize}
    \item Kid: push-recovery $\rightarrow$ blindfolded robot challenge
    \item High jump $\rightarrow$ Parkour challenge
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Planned for 2020, but canceled}
  \begin{itemize}
  \item Humanoid Open Competition
  \item Technical Challenges
    \begin{itemize}
    \item Kick from moving ball: enforce that pass is performed by a robot
    \item Adult: push-recovery $\rightarrow$ standing-up
    \end{itemize}
  \item Discussion: How do we manage the upcoming changes?
    \begin{itemize}
    \item Mandatory league vote before adding something
    \item Can we revote on something that has been announced
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Planned for 2021}
  \begin{itemize}
  \item Being more explicit on network requirements
    \begin{itemize}
    \item WiFi quality: specification of a minimal bandwidth, anywhere on the field
    \end{itemize}
  \item Enforcing the rules more strictly
    \begin{itemize}
    \item Monitoring bandwidth of each team
    \item Robots not in play should not be able to send messages
    \end{itemize}
  \item Discussion:
    \begin{itemize}
    \item Require significant development (any volunteer)
    \item Risk that we set rules which can not be applied
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Discussion: Height limits}
  \begin{itemize}
  \item Concerns expressed
    \begin{itemize}
    \item Adult: Lowering min height is taking a step back
    \item Kid: Having 40cm and 1.2m high robots in the same league is difficult
    \end{itemize}
  \item This rule has major impact on both leagues, how can we vote on that?
    \begin{itemize}
    \item More teams in Kid: unfair for Adult
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Discussion: Next major update}
  \begin{itemize}
  \item 2023 or 2024?
  \item What content?
  \end{itemize}
\end{frame}

\section{Long term: event-triggered roadmap}

\begin{frame}{Metrics}
  \begin{block}{Motion metrics}
    Peak speed, ball approach duration, Kick properties (distance, accuracy, ...)
  \end{block}
  \begin{block}{Perception metrics}
    Ball-localization, self-localization, opponent recognition
  \end{block}
  \begin{block}{Gameplay metrics}
    Shots on goal, diving saves, nb robots on fields, Don't Mess Up Period (DMUP)
  \end{block}
  \begin{block}{Requirements}
  Compare \textbf{ground truth} (labeled video) with \textbf{robots
    prediction} (common protocol).
  \end{block}
\end{frame}

\begin{frame}{About criteria and rule implementation}
  \begin{block}{Shared success}
    At least \textbf{3 teams} need to fit a criterion to consider that the step
    is completed.
  \end{block}
  \begin{block}{Time to plan for participants and LOC}
    When a step is completed, the associated change is triggered
    \textbf{two years later}.
  \end{block}
\end{frame}
%TODO add note that 3 teams need to reach criteria
%TODO add note that there is a two years deadline between completion of the
%criteria, and the effective change

\begin{frame}{Field dimensions and number of robots}
  \begin{block}{Evolution planned}
    \centering
    \begin{tabular}{c|r c r|r c r}
      \textbf{Robots} & \multicolumn{3}{c}{\textbf{Kid}} & \multicolumn{3}{c}{\textbf{Adult}}\\
      \hline
      2 & 4.5 & $\times$ & 3 & \textbf{14} & $\times$ & \textbf{9}\\
      3 & 6 & $\times$ & 4 & 21 & $\times$ & 14\\   
      4 & \textbf{9} & $\times$ & \textbf{6} & 32 & $\times$ & 21\\
      6 & 14 & $\times$ & 9 & 43 & $\times$ & 32\\
      8 & 21 & $\times$ & 14 & 65 & $\times$ & 43\\
      11 & 32 & $\times$ & 21 & 105 & $\times$ & 70
    \end{tabular}
  \end{block}
  \begin{block}{When is size increased?}
    \begin{itemize}
    \item Peak speed: 20s to cross the field.
    \item Kick distance: at least half a field.
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Discussion: Outdoor?}
  \begin{itemize}
  \item Critical to increase field dimensions and number of players
  \item Gradual steps
    \begin{itemize}
    \item Outdoor with a roof first
    \end{itemize}
  \item Moving from artificial turf to real grass? Is it necessary?
  \item When/how can we plan it? 
    \begin{itemize}
    \item Include preliminary testing through technical challenges?
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Handlers and game duration}
  \begin{block}{Adult: path to remove the handlers}
    \begin{enumerate}
    \item TC: Stand-Up (less than 10 seconds)
    \item TC: Falling on hard ground
    \end{enumerate}
    Once robot can properly fall and stand-up, handlers will be removed
  \end{block}

  \begin{block}{Towards autonomy}
    \begin{itemize}
    \item Once DMUP > 3 min: harsher penalties for pick-up and service
    \item When $\text{DMUP} > \frac{\text{halfTimeDuration}}{2}$: increase half
      time duration:
      \begin{itemize}
      \item Steps: 15, 20, 30, 45 minutes.
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Gameplay rules}
  \begin{block}{Mandatory throw-in}
    \begin{itemize}
    \item Throw-in takes less than 10 seconds to be performed
    \item Ball cross at least quarter of field length, before stopping
    \end{itemize}
  \end{block}
  \begin{block}{Offside rule}
    \begin{itemize}
    \item Satisfying self and opponent localization ($\text{RMSE} < \frac{\text{fieldLength}}{50}$)
    \item At least 6 robots per team
    \end{itemize}
  \end{block}
\end{frame}


\end{document}
